---
title: "Day 5 MORE Review for Quiz"
author: "Stephanie Neul"
date: "2022-09-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

1. Read in the data set ```anothersurvey.csv``` and display the first *8* rows. This data came from this survey - feel free to look at the questions asked in case it's helpful for any of these questions:
https://docs.google.com/forms/d/e/1FAIpQLSeTE0TN6CykbqhVmQJSlDhWNxu2XVGGlNUX6Gu5GdhoBiQ4GQ/viewform?usp=sf_link

```{r}
data.set = read.csv("anothersurvey.csv")
head(data.set, 8)
```

2. How many people filled out the survey? Support your answer with code, don't just count :)

```{r}
nrow(data.set)
```

3. How many variables are in this data frame?

```{r}
ncol(data.set)
```

4. Explain why the variable `form` should be a factor, then change it to be a factor.

```{r}
# Form should be a factor because there's only four options: 3rd, 4th, 5th, or 6th. Also, the same form means the same thing for one person as it does for another as in it's not unique.

data.set$form = as.factor(data.set$form)
```


5. Choose a variable that you believe should be turned into a logical vector. Explain why, and then change it into a logical vector. There a several possible options, just choose one.

```{r}
# Are you a PG should be turned into a logical vector because there are only two cases, yes or no.

data.set$are.you.a.PG = data.set$are.you.a.PG == "Yes"
data.set$are.you.a.PG
```

6. Respondents were asked to rate "How neat is your room" on a scale of 0-5. It's more typical to use a scale from 1-5, though, so let's make all 0's into 1's. Change the `how.neat.is.your.room vector` so that any answers of 0 are now 1.

```{r}
data.set$how.neat.is.your.room[data.set$how.neat.is.your.room == 0] = 1
data.set$how.neat.is.your.room
```

7. Make a new column in the data frame called `home.km.away.from.choate` that takes the response to `home.miles.away.from.choate` and converts it into kilometers. If you don't know how many kilometers are in a mile, look it up :)

```{r}
data.set$home.km.away.from.choate = data.set$home.miles.away.from.choate * 1.609
data.set$home.km.away.from.choate
```

8. Two people answered that their middle names were "n/a". A more descriptive answer would be "No middle name". Change those answers to "No middle name".

```{r}
data.set$middle.name[data.set$middle.name == "n/a"] = "No middle name"
data.set$middle.name
```

9. Use sorting and indexing to print out the greatest number of miles from Choate to a survey respondent's home.

```{r}
sort(data.set$home.miles.away.from.choate, TRUE)[1]
```

10. Unless you already did this in problem 5, change `do.you.like.veg.out.monday` into a logical vector. Note that you will have to make a choice when doing this; explain why you made the choice that you did.

```{r}
data.set$do.you.like.veg.out.monday = data.set$do.you.like.veg.out.monday == "Yes"
data.set$do.you.like.veg.out.monday

# I think anything short of a complete yes can't be considered true. Just like half-truths don't actually count as a truth.
```

11. Print out all of the distances from Choate to home that are less than 100 miles.

```{r}
data.set$home.miles.away.from.choate[data.set$home.miles.away.from.choate < 100]
```

12. Is `middle.name` a categorical or quantitative variable? Also, what class is it currently saved as? Is that the best choice of class for this variable?

```{r}
# middle.name is a categorical variable, it fails to past the "average" test. Besides, middle names are used as identifiers, not as measurements. middle.name is currently saved as character. I think it is the best choice of class for this variable.

class(data.set$middle.name)
```

