---
output:
  pdf_document: default
  html_document: default
---
8. a.
k = 6, n = 100, p = 0.04
```{r}
1 - pbinom(5, 100, 0.04)
```
8. b.
p = 0.04
```{r}
sqrt(0.04*0.96/100)
```
sd = 0.0196
```{r}
1 - pnorm(0.06, 0.04, 0.0196)
```
8. c.
k = 57, n = 100, p = 0.51
```{r}
1 - pbinom(56, 100, 0.51)
```
8. d.
p = 0.51
```{r}
sqrt(0.51*0.49/100)
```
sd = 0.05
```{r}
1 - pnorm(0.57, 0.51, 0.05)
```
8. e. The percent difference between the answers to a) and b), in which np and nq are less than 10, is greater than the percent difference between the answers to c) and d), in which np and nq are greater than 10. np and nq being greater than 10 results in more similar Binomial and Normal distributions.

21. a.
p = 0.07
```{r}
sqrt(0.07*0.93/200)
```
sd = 0.018
21. b.
Bernoulli trials and np and nq are greater than 10

Bernoulli trials:
success/failure: yes; success = don't pay back loan on time; failure = pay back loan on time
independent: no, drawing without replacement
OR
200 < 10% of all borrowers; most likely true
constant probability of success: yes because the 200 are drawn at random

np >= 10:
200 * 0.07 = 14
nq >= 10:
200 * 0.93 = 186
21. c.
```{r}
1 - pnorm(0.1, 0.07, 0.018)
```

22. a.
Both the Binomial and Normal distributions are appropriate.

We have Bernoulli trials:
success/failure: yes; success = use phone to connect; failure = does not use phone to connect
independence: no, drawing without replacement
OR
100 < 10% of all teens, yes
constant probability of success: yes, drawn at random

np >= 10:
100 * 0.84 = 84
nq >= 10:
100 * 0.16 = 16

p = 0.84
```{r}
sqrt(0.84*0.16/100)
```
sd = 0.037

22. b.
```{r}
1 - pnorm(0.9, 0.84, 0.037)
```