---
title: "Day 5 Review for Quiz Solutions"
author: "Stephanie Neul"
date: "9/20/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

##Part A 

1. Find the cubed root of 30.
```{r}
30^(1/3)
```

2. Create a character vector of the first names of four people in this class, then display your vector.

```{r}
d.block.names.vector <- c("Naomi", "Adriana", "Tom", "Jack")
d.block.names.vector
```

3. Have R display your character vector from problem 2 so that it is sorted in alphabetical order (from A to Z).

```{r}
sort(d.block.names.vector, decreasing = FALSE)
```

4. A teacher has given a test. The following are the scores from one of her classes. Create a vector of the scores and then sort the vector so that the scores are in order from highest to lowest score. Test scores: 61, 100, 98, 85, 75, 87, 90, 101, 99, 80, 60, 90, 101, 65

```{r}
test.scores <- c(61, 100, 98, 85, 75, 87, 90, 101, 99, 80, 60, 90, 101, 65)
sort(test.scores, decreasing = TRUE)
```


  a. The teacher has decided that anyone who scored lower than 70 will automatically earn a score of 70. Use a logical test to make this change.


```{r}
test.scores[test.scores < 70] <- 70

test.scores
```

  b. The teacher has decided to also add 5 points to everyone's test score. Create a *new* vector that contains the new scores.

```{r}
new.test.scores <- test.scores + 5
new.test.scores
```


5. Display every integer below 100 that is divisible by 7.


```{r}
seq(from = 0, to = 100, by = 7)
```


6. List all of the integers from 60 to 72.

```{r}
60:72
```

##Part B

1. Read in the data set ```nba_players_19.csv``` and display the first *7* rows. You can find more information about the data set here: https://www.openintro.org/data/index.php?data=nba_players_19

```{r}
nba.data <- read.csv("nba_players_19.csv")
head(nba.data, 7)
```


2.  How many observations and variables does this data frame have? How do you know (You must use R to find these values)?

```{r}
nrow(nba.data)
ncol(nba.data)
```

494 observations, 7 variables

3. For the variable *last_name*, answer the following questions:

a. What type of variable? (Quantitative or categorical)

Categorical


b. What class should it be, and why?

Character, because last names are unique and do not have categories. Each last name should be treated as a unique element.

4. For the variable *team*, answer the following questions:

a. What type of variable? (Quantitative or categorical)

Categorical


b. What class should it be, and why?

Factor, because there are a finite number of categories for this variable, and we would like R to recognize that, for example, the value "Nets" for team name manes the same thing for everyone who has that value for team name.

5. For the variable *number*, answer the following questions:

a. What type of variable? (Quantitative or categorical)

Categorical (it's like an id number or label! Number = jersey number)


b. What class should it be, and why?

Character or numeric or factor (could make an argument for any of those, there is not one right answer)

6. For the variable *height*, answer the following questions:

a. What type of variable? (Quantitative or categorical)

Quantitative


b. What class should it be, and why?

Numerical (has units!)

7. For each variable that you said above is not the correct class, change it in the data frame so that it is the correct class. Then display the first 7 rows of the updated data frame.

```{r}
nba.data$number <- as.character(nba.data$number) #I thought it should be character, personally, but you could leave it as numeric

nba.data$team <- as.factor(nba.data$team)

head(nba.data, 7)
```


8.Change the position vector to a logical vector that is TRUE if the player is a Guard and FALSE otherwise.
```{r}
nba.data$position <- nba.data$position == "Guard"

```




9. This data frame gives the height of each player in inches. Create a new vector in the data frame that gives the height of the player in centimeters. (1 inch = 2.54 cm). Then display the first 7 rows of the updated data frame.

```{r}
nba.data$height.in.cm <- nba.data$height * 2.54

head(nba.data, 7)
```


10. What is the height in inches of the tallest player in this data set? Use sorting and element extraction for this problem.

```{r}
sorted.heights <- sort(nba.data$height, decreasing = TRUE)
sorted.heights[1]
```


11. Print out the heights of all players in this data set who are over 7 feet tall.

```{r}
nba.data$height[nba.data$height > 84]
```

