---
title: "Day 23 Intro to Probability E Block"
author: "Stephanie Neul"
date: "11/07/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

You are going to write a function to simulate a very simple probability problem. We will use the function to explore the concept of Empirical Probability as well as the Law of Large Numbers.

Your task:
Write a function called coin.flipper() that simulates flipping a fair coin a specified number of times. It should output the *proportion of flips* that were heads.

Before you start, answer the following questions:
1. What does the function input? What does it output?

Input number of times to flip. Output proportion of flips that are heads.

2. What do you need to keep track of? Choose names for those variables.

Total number of flips and heads.

3. How exactly will you simulate flipping a coin (i.e. how will you use sample, how will you see if it's heads or tails)

sample(1:2, 1)

4. What is the condition for breaking the loop?
if number of times we have flipped is the specified number to flip
```{r}
coin.flipper <- function(number.to.flip){
    #initialize variables
    heads = 0
    coins.flipped = 0
    
    repeat {
        flip = sample(1:2, 1)
        
        if (flip == 1) {
            heads = heads + 1
        }
        
        coins.flipped = coins.flipped + 1
        
        if (coins.flipped >= number.to.flip) {
            break
        }
    }
    
    return(heads/coins.flipped)
}
```

Test your function:
```{r}
coin.flipper(10000000)
```

Replicate your function 1000 times, with an input of *100 flips* for each trial. Save the results in a vector.
```{r}
thousand.trials.hundred.flips = replicate(1000, coin.flipper(100))
```

Histogram for 1000 replications of 100 flips:
```{r}
hist(thousand.trials.hundred.flips)
```

Replicate your function 1000 times, this time with an input of *1000 flips* for each trial. Again save the results in a vector.
```{r}
summary(thousand.trials.hundred.flips)
```

Histogram for 1000 replications of 1000 flips:
```{r}
thousand.trials.thousand.flips = replicate(1000, coin.flipper(1000))
hist(thousand.trials.thousand.flips)
```

## The law of large numbers:
The law of large numbers states that as the number of independent trials increases, the *long-run* **relative frequency** of repeated events gets closer and closer to a single value.