---
title: "Day 08 Boxplots in R"
author: "Carey Kopeikin and Matt Bardoe edited by Stephanie Neul"
date: "9/22/2021"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

# What you will learn

-   How to graph boxplots using R
-   Parameters for formatting graphs in R

#### Load the Data

Let's again load the data `avocado_updated_2020_total_us.csv`. From the
creator of the data set:

"It is a well known fact that Millennials LOVE Avocado Toast. It's also a well known fact that all Millennials live in their parents basements. Clearly, they aren't buying homes because they are buying too much Avocado Toast! But maybe there's hope... if a Millennial could find a place with cheap avocados, they could live out the Millennial American Dream."

**Read about the data before you get started:** This data set is an
updated version of the one described here:
<https://www.kaggle.com/neuromusic/avocado-prices?select=avocado.csv>
The updates are described here; most importantly, the updated dataset
now includes data up through November 2020:
<https://www.kaggle.com/datasets/timmate/avocado-prices-2020>

And again, to start, we are going to look at data for the US as a whole,
for "conventional" (non-organic) avocados only. Each row contains the
sales data for the entire US for one week.

```{r}
avocado <- read.csv("avocado_updated_2020_total_us.csv")
head(avocado)
```

#### Boxplots in R

Last class you learned how to draw a boxplot by hand. Moving forward, we
are going to have R draw our box plots for us.

As a reminder: In a boxplot a box is drawn between Q1 and Q3, and there
is a line in the box to reflect the median. Then lines are drawn to
connect the max to end of the box at Q3 and another to connect the min
to the box at Q1. Any outliers are drawn as dots, and in that case the
largest and smallest non-outlier points are connected to the box
instead.

Here is how you would use R to create a boring boxplot for the variable
`average_price`.

```{r}
boxplot(avocado$average_price)
```

*Describe or discuss what this boxplot shows us*

Unfortunately this graph is not visually appealing. Here are some ways
you can change your graph:

+-----------+------------------------------------------------+
| Parameter | Definition                                     |
+===========+================================================+
| *type*    | Tells R how to plot the points joined with     |
|           | lines or individually                          |
+-----------+------------------------------------------------+
| *main*    | The main title for the graph                   |
+-----------+------------------------------------------------+
| *xlab*    | The label for the x-axis                       |
+-----------+------------------------------------------------+
| *ylab*    | The label for the y-axis                       |
+-----------+------------------------------------------------+
| *col*     | The color or colors to use for plotting        |
+-----------+------------------------------------------------+
| *pch*     | The point character. This selects the          |
|           | individual characters for points               |
+-----------+------------------------------------------------+
| *cex*     | controls the size of the points                |
+-----------+------------------------------------------------+
| *lty*     | The line type                                  |
+-----------+------------------------------------------------+
| *lwd*     | The line width                                 |
+-----------+------------------------------------------------+
| *xlim*    | A vector that provides the limits for the      |
|           | horizontal axis                                |
+-----------+------------------------------------------------+
| *ylim*    | A vector that provides the limits for the      |
|           | vertical axis                                  |
+-----------+------------------------------------------------+

Note that not all of these are used for all graphs. To find out more
type the kind of plot you want to learn about in the help search or

```{r}
?boxplot
```

You can read more about all the point styles here:
<http://www.sthda.com/english/wiki/r-plot-pch-symbols-the-different-point-shapes-available-in-r>

You can see more about colors here:
<http://www.sthda.com/english/wiki/colors-in-r>

Now here is a less-boring boxplot:

```{r}
boxplot(avocado$average_price,
        main = "Average Weekly Prices of Avocado in the US, 2016-2020",
        ylab = "US Dollars",
        col = "lightgreen",
        border = "blue",
        outcol = "red"
        )

```

#### Comparative Box Plots

We can also make comparative box plots - which means several box plots
side by side. The one below shows the average price of avocados by year.

```{r}
boxplot(avocado$average_price~avocado$year,
        main = "Average Prices of Avocados by Year",
        ylab = "US Dollars", 
        xlab = "Year", 
        col = c("green", "blue"),
        border = "black",
        outcol = "red")
```

We can make each box a different color if we want! You just need to add
colors to the col vector. Try it!

We can also turn the axis labels sideways by setting the parameter `las` equal to 2, and change the size of the axis labels: 

```{r}
boxplot(avocado$average_price~avocado$year,
        main = "Average Prices of Avocados by Year",
        ylab = "US Dollars", 
        xlab = "Year", 
        col = c("green", "blue"),
        border = "black",
        outcol = "red",
        las = 2, #change orientation of axis labels
        par(cex.axis = .6) #size of axis labels
      )
```

It is said that there was an "avocadopocalypse" in 2017. Explain what
you think people meant by that, based on the graph above.

**Your Turn** Write code to make a boxplot of Total Volume by Year. Feel
free to copy and paste code from above and change what you need to
change. Make sure the graph is labeled appropriately!

```{r}
boxplot(avocado$total_volume~avocado$year,
        main = "Total Volume by Year",
        ylab = "US Dollars", 
        xlab = "Year", 
        col = c("green", "blue"),
        border = "black",
        outcol = "red",
        las = 2, #change orientation of axis labels
        par(cex.axis = .6) #size of axis labels
      )
```

According to this graph, did the "avocadopocalypse" seem to impact
volume of avocados sold?

#### Subsetting a Dataframe

Sometimes we may just want to look at one particular year. To do so, we
can make a *new dataframe* that only contains data for that year, using
the following code. This is called *subsetting* the dataframe. You can
do this anytime you want to separate out the data for just a particular
category. The code below makes a new data frame containing only data for
avocados in 2018.

**DO NOT FORGET THE COMMA!!!**

```{r}
avocado.2018 <- avocado[ avocado$year == 2018, ]
head(avocado.2018)
```

Now, we can calculate the average weekly volume and the weekly average
price of avocados sold for just 2018:

```{r}

```

And we can make a boxplot of average prices for just 2018:

```{r}
boxplot(avocado.2018$average_price,
        main = "Average Weekly Prices of Avocados in 2018",
        ylab = "US Dollars",
        col = "lightgreen",
        border = "blue",
        outcol = "red"
        )
```

Make some observations about the boxplot above:

We may be interested in learning more about the outliers we see. What
week were avocados most expensive, and least expensive? We can order the
data to find out:

```{r}
avocado.2018[ order(avocado.2018$average_price, decreasing = TRUE),]

```

**Your Turn** Hass Avocado splits the US into 8 regions for purposes of
sales data collection: California, Great Lakes, Midsouth, Northeast,
Southeast, South Central, Plains, and West. They also track organic
avocado sales.

Read in the dataset called `avocado_updated_2020_large_regions.csv`. It
contains data for each of the regions, and for both conventional and
organic avocados.

```{r}
avocado.regions <- read.csv("avocado_updated_2020_large_regions.csv")
```

1.  Make a comparative boxplot for the average price of avocados by
    `type`. Make some observations about your boxplot.

2.  Make a new dataframe by subsetting that contains the data only for
    organic avocados. Then make a boxplot for the average price of just
    organic avocados.

3.  Where would you recommend that I live if I want to have access to
    the cheapest avocados? Explain, and make sure to use a graph to
    support your answer. Make sure the labels on your graph are properly formatted!!!

4.  Does your answer to question 3 change if I tell you I will only buy
    organic avocados? Explain.

5.  Which region bought more avocados in 2019: California or the
    Northeast? Hint: you shouldn't need to make a graph to answer this
    question.

6.  Bonus: Make a comparative boxplot that shows the amount of money
    spent on avocados by year in the US as a whole. You'll have to add a
    column to the dataframe.
